### 2025-06-04 paper/intro.tex
Rewrite introduction.

The year is 2025, and we still have to make a formidable choice between
correctness and performance for all the programming projects that we start.
Low level programming languages such as C or Fortran make it possible to leverage
intricate hardware features for the price of poor analysability and
correctness guarantees.  Dependently-typed systems such as Lean or Agda make it possible
to describe arbitrary invariants within the given program, yet they can
rarely generate high-performance code.

The idealist inside of us is exasperated, because there should be a
perfect solution that caters for both cases. However, the practitioner
within us proposes a different perspective. Instead of using a single
language, we may use two languages in cooperation. Specifically, we
envision an alliance between a proof assistant and some
high-performance language of our choice, where we extract code from
the latter to the former.

@@

@@

As a result, we demonstrate an approach where the entire specification,
optimiser, AD and code generation are available to us within a proof
assistant of our choice.  We can prove facts about all the
stages of the pipeline and easily adjust them to our liking.
We argue that such a liberating
approach is feasible in practice, at least for the times of dialectic
of correctness and performance.

>>>>

The year is 2025, and programmers still face a trade-off between correctness and
performance. Low-level languages like C or Fortran allow careful control over
hardware, but offer few guarantees about safety or program structure.
Dependently-typed languages like Agda or Lean, on the other hand, make it
possible to encode precise invariants, but typically lack the infrastructure for
high-performance computation.

Rather than chasing a mythical language that promises both complete correctness
and high performance, we explore a practical compromise: leveraging a
dependently-typed proof assistant for the parts of a scientific programming
pipeline where full verification is most crucial, and integrating it with a
high-performance backend for execution. Specifically, we investigate how Agda
can be used not only for specification and reasoning, but also for
transformation, optimisation, and code generation in a realistic machine
learning workload.

@@

In an ideal world, all parts would of course come with full formal
specifications and accompanying proof, but in practice some parts remain an
entirely separate and nontrivial research effort to verify for functional
correctness, such as AD. Other parts---such as code generation--can be handled
using known techniques, but are somewhat time consuming. Despite all this, we
show that even in those cases where resources preclude full verification, Agda
can still be used as a practical tool, with full verification employed where it
is deemed desirable and practical.

@@

As a result, we demonstrate an approach where the entire specification,
optimiser, AD and code generation are available to us within a proof assistant
of our choice. We can prove facts about all the stages of the pipeline and
easily adjust them to our liking. We also demonstrate that some components can
be left only partially verified, where the effort of verification is judged
disproportionate to the benefit. We argue that such a liberating approach is
feasible in practice, at least for the times of dialectic of correctness and
performance.
